---
title: "Secretome analysis"
author: "Leon Heffinger"
date: "2025-11-13"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,        # TRUE = Show code by default; FALSE = Hide code by default
  message = FALSE,     # Suppress messages
  warning = FALSE,     # Suppress warnings
  fig.align = "center",
  fig.width = 32,
  fig.height = 30
)

```

```{r install.pkg, echo=FALSE, message=FALSE, warning=FALSE, include=TRUE}
install.pkg <- function(x){
  new.pkg <- x[!(x %in% installed.packages()[, "Package"])]
  if (length(new.pkg) > 0){
    install.packages(new.pkg, dependencies = TRUE)
    sapply(x, require, character.only = TRUE)
  }else{
    sapply(x, require, character.only = TRUE)
  }
}
packages <- c("broom", "dplyr", "BiocManager", "remotes", "ggpubr", "tibble", 'tidyverse', 'ggplot2', 'devtools', 'factoextra', 'stringr', "magrittr" , "circlize", "NMF", "bsicons", "parallel", "grDevices", "colorRamps","RColorBrewer", "ggalluvial", "readr", "umap", "sf", "spatialreg","spatstat", "readxl","drc", "Rmisc", "tidyr", "purrr", "pheatmap", "RColorBrewer", "scales", "colorspace", "nortest", "rstatix", "broom", "dplyr", "car", "ggpubr", "tibble", 'stringr', "pheatmap", "reshape2", "ggpattern", "gtsummary", "gt", "rmarkdown")
install.pkg(packages) 
```

## Functions

```{r}
is_outlier <- function(x, threshold = 0.75) {
  q <- quantile(x, probs = c(0.25, 0.75))
  iqr <- diff(q)
  lower_bound <- q[1] - threshold * iqr
  upper_bound <- q[2] + threshold * iqr
  outliers <- x < lower_bound | x > upper_bound
  return(outliers)
}
  
normalize_columns_by_group <- function(df, id_column, columns_to_normalize, reference_column) {
  # Group by the ID column and normalize the specified columns
  df <- df %>%
    group_by(!!sym(id_column)) %>%
    mutate(across(all_of(columns_to_normalize), 
                  ~ . / mean(get(reference_column), na.rm = TRUE), 
                  .names = "{.col}_normalized")) %>%
    ungroup()
  
  return(df)
}

#function for mean calculation and putting sds into plots
data_summary <- function(x) {
  m <- mean(x)
  ymin <- m-sd(x)
  ymax <- m+sd(x)
  return(c(y=m,ymin=ymin,ymax=ymax))
}
```

# Data Import

```{r}
secretome_data<- read_excel("Secretome_data.xlsx", sheet = 1)%>%
  mutate(Description_base = tools::file_path_sans_ext(basename(Description))) %>%
  separate(Description_base, into = paste0("p", 1:2), sep = "_", fill = "right", remove = FALSE) %>%
  transmute(
    across(everything()),
    Lab_number = p1,          
    Location      = p2,          
  )
secretome_data <- secretome_data[ , !grepl("^p\\d+$", names(secretome_data))]

secretome_data$Location[secretome_data$Location == "C"] <- "Central"
secretome_data$Location[secretome_data$Location == "P"] <- "Peripheral"
secretome_data$Location[secretome_data$Location == "NT"] <- "None_tumor"

patient_data<- read_excel("Secretome_data.xlsx", sheet = 2)


#join on column "id" present in both data frames
secretome_data_merged <- secretome_data %>%
  left_join(patient_data, by = "Lab_number")
secretome_data_merged$comment <- NULL
secretome_data_merged <- na.omit(secretome_data_merged)
colnames(secretome_data_merged)[48] <- "Sex"


cytokines <- c("Hu_IL_2_(38)", "Hu_IL_8_(54)", "Hu_IL_10_(56)", "Hu_IFN_g_(21)", "Hu_Eotaxin_(43)", "Hu_IP_10_(48)", "Hu_MIP_1a_(55)", "Hu_IL_16_(27)","Hu_CTACK_(72)","Hu_GROa_(61)","Hu_MCP_3_(26)","Hu_MIF_(35)" ,"Hu_MIG_(14)","Hu_SDF_1a_(22)","Hu_IL_1b_(39)", "Hu_IL_4_(52)","Hu_IL_6_(19)","Hu_GM_CSF_(34)","Hu_MCP_1(MCAF)_(53)", "Hu_TNF_a_(36)", "CCl21_(12)","CXCL13_(74)", "CXCL5_(73)","CCL24_(30)","CCl26_(65)","CX3CL1_(77)", "CXCL6_(15)", "CXCL2_(78)", "CCl1_(20)", "CXCL11_(25)", "CCL8_(57)", "CCL13_(28)", "CCL22_(29)", "CCL15_(66)", "CCl20_(62)", "CCL19_(76)", "CCL23_(37)", "CXCL16_(64)", "CCl17_(67)", "CCL25_(46)")  # adjust

rename_cytokines <- c("IL2", "IL8", "IL10", "IFNG", "EOTAXIN", "IP10", "MIP1A", "IL16", "CTACK", "GROA", "MCP3", "MIF", "MIG", "SDF1A", "IL1B", "IL4", "IL6", "GMCSF", "MCP1", "TNFA", "CCL21", "CXCL13", "CXCL5", "CCL24", "CCL26", "CX3CL1", "CXCL6", "CXCL2", "CCL1", "CXCL11", "CCL8", "CCL13", "CCL22", "CCL15", "CCL20", "CCL19", "CCL23", "CXCL16", "CCL17", "CCL25")

 #Find overlaps between df columns and old_names
to_change <- intersect(cytokines, names(secretome_data_merged))# Replace only those, preserving the correct mapping
idx_old <- match(to_change, names(secretome_data_merged))
idx_new <- match(to_change, cytokines)
names(secretome_data_merged)[idx_old] <- rename_cytokines[idx_new]# Optional: make names unique if any collisions occurnames(df) <- make.unique(names(df))
```

### Filtering healthy donors and sex

```{r}
secretome_data_merged_PDAC <- secretome_data_merged %>%
  filter((`PC_(no_0_yes__1)` == 1))
secretome_data_merged_PDAC <- secretome_data_merged_PDAC %>%
  filter(Diagnosis_for_display_purposes == "PDAC")
secretome_data_merged_PDAC_females <- secretome_data_merged_PDAC %>%
  filter((Sex == 1))
secretome_data_merged_PDAC_males <- secretome_data_merged_PDAC %>%
  filter((Sex == 0))
secretome_data_merged_healthy_donors <- secretome_data_merged %>%
  filter((`PC_(no_0_yes__1)` == 0))

```

```{r}

secretome_data_merged_PDAC_females_long <- secretome_data_merged_PDAC_females %>%
  pivot_longer(cols = all_of(rename_cytokines), names_to = "Cytokine", values_to = "Expression") %>%
  group_by(Location, Lab_number, Cytokine, Description)
secretome_data_merged_PDAC_males_long <- secretome_data_merged_PDAC_males %>%
  pivot_longer(cols = all_of(rename_cytokines), names_to = "Cytokine", values_to = "Expression") %>%
  group_by(Location, Lab_number, Cytokine, Description)
secretome_data_merged_long_boxplots <- rbind(secretome_data_merged_PDAC_males_long ,secretome_data_merged_PDAC_females_long )

secretome_data_merged_long_boxplots <- secretome_data_merged_long_boxplots %>%  
  group_by(Lab_number, Cytokine) %>%  
  mutate(    baseline = Expression[Location == "Plasma"],  # Plasma baseline for each patient & cytokine    
             FoldChange = Expression / baseline,            # FC relative to Plasma
             Log2_FC = log2(FoldChange)
             ) #%>%  ungroup() %>%  
              #select(-baseline)  # Optional: remove baseline column if not needed# Check result

secretome_data_merged_long_boxplots$Sex <- as.character(secretome_data_merged_long_boxplots$Sex)
secretome_data_merged_long_boxplots['Sex'][secretome_data_merged_long_boxplots['Sex'] == '0'] <- 'Male'
secretome_data_merged_long_boxplots['Sex'][secretome_data_merged_long_boxplots['Sex'] == '1'] <- 'Female'

secretome_data_merged_PDAC_females_long_summary <- summarySE(secretome_data_merged_PDAC_females_long, measurevar="Expression", groupvars=c("Location", "Cytokine"))
secretome_data_merged_PDAC_females_long_summary['Sex'] = 'Female'

secretome_data_merged_PDAC_males_long_summary <- summarySE(secretome_data_merged_PDAC_males_long, measurevar="Expression", groupvars=c("Location", "Cytokine"))
secretome_data_merged_PDAC_males_long_summary['Sex'] = 'Male'
```

## Calculations for heatmaps where the baseline is Non_tumor instead of Plasma

```{r}
secretome_data_merged_long_boxplots_none_tumor_baseline <- rbind(secretome_data_merged_PDAC_males_long,secretome_data_merged_PDAC_females_long )

secretome_data_merged_long_boxplots_none_tumor_baseline <- secretome_data_merged_long_boxplots_none_tumor_baseline %>%  
  group_by(Lab_number, Cytokine) %>%  
  mutate(    baseline = Expression[Location == "None_tumor"],  # Plasma baseline for each patient & cytokine    
             FoldChange = Expression / baseline,            # FC relative to Plasma
             Log2_FC = log2(FoldChange)
             ) #%>%  ungroup() %>%  
              #select(-baseline)  # Optional: remove baseline column if not needed# Check result

secretome_data_merged_long_boxplots_none_tumor_baseline$Sex <- as.character(secretome_data_merged_long_boxplots_none_tumor_baseline$Sex)
secretome_data_merged_long_boxplots_none_tumor_baseline['Sex'][secretome_data_merged_long_boxplots_none_tumor_baseline['Sex'] == '0'] <- 'Male'
secretome_data_merged_long_boxplots_none_tumor_baseline['Sex'][secretome_data_merged_long_boxplots_none_tumor_baseline['Sex'] == '1'] <- 'Female'

secretome_data_merged_PDAC_females_long_summary <- summarySE(secretome_data_merged_PDAC_females_long, measurevar="Expression", groupvars=c("Location", "Cytokine"))
secretome_data_merged_PDAC_females_long_summary['Sex'] = 'Female'

secretome_data_merged_PDAC_males_long_summary <- summarySE(secretome_data_merged_PDAC_males_long, measurevar="Expression", groupvars=c("Location", "Cytokine"))
secretome_data_merged_PDAC_males_long_summary['Sex'] = 'Male'
```

## Boxplots Cytokines Central

```{r}
secretome_data_merged_long_boxplots_central<- secretome_data_merged_long_boxplots %>%
  filter((Location == "Central"))

Boxplots_central <- ggplot(secretome_data_merged_long_boxplots_central, aes(x = Sex, y = Expression, fill = Sex)) +
  geom_boxplot(outlier.shape = NA, width = 0.65, alpha = 0.85) +
  stat_compare_means(
    aes(group = Sex),
    method = "wilcox.test", # or "t.test" depending on your analysis
    paired = FALSE,
    label = "p.signif",     # shows stars instead of p-values
    label.x.npc = "center"
  ) +
  geom_jitter(aes(color = Sex),
              width = 0.15, height = 0, size = 1.6, alpha = 0.5, show.legend = FALSE) +
  facet_wrap(~ Cytokine, scales = "free_y") +
  scale_fill_manual(values = c(Male = "#0000FF", Female = "#FF0000")) +
  scale_color_manual(values = c(Male = "#000080", Female = "#D50000")) +
  labs(
    x = "Sex",
    y = "Expression",
    title = "Cytokine Expression by Sex Central",
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.spacing = unit(1, "lines"),
    strip.text = element_text(face = "bold"),
    legend.position = "bottom"
  )

#Print the plot
print(Boxplots_central)

#Save to file
ggsave("cytokine_boxplots_Central.pdf", Boxplots_central, width = 18, height = 13)

```

## Boxplots Cytokines Peripheral

```{r}
secretome_data_merged_long_boxplots_Peripheral<- secretome_data_merged_long_boxplots %>%
  filter((Location == "Peripheral"))

Boxplots_Peripheral <- ggplot(secretome_data_merged_long_boxplots_Peripheral, aes(x = Sex, y = Expression, fill = Sex)) +
  geom_boxplot(outlier.shape = NA, width = 0.65, alpha = 0.85) +
  stat_compare_means(
    aes(group = Sex),
    method = "wilcox.test", # or "t.test" depending on your analysis
    label = "p.signif",     # shows stars instead of p-values
    label.x.npc = "center"
  ) +
  geom_jitter(aes(color = Sex),
              width = 0.15, height = 0, size = 1.6, alpha = 0.5, show.legend = FALSE) +
  facet_wrap(~ Cytokine, scales = "free_y") +
  scale_fill_manual(values = c(Male = "#0000FF", Female = "#FF0000")) +
  scale_color_manual(values = c(Male = "#000080", Female = "#D50000")) +
  labs(
    x = "Sex",
    y = "Expression",
    title = "Cytokine Expression by Sex Peripheral",
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.spacing = unit(1, "lines"),
    strip.text = element_text(face = "bold"),
    legend.position = "bottom"
  )

#Print the plot
print(Boxplots_Peripheral)

#Save to file
ggsave("cytokine_boxplots_Peripheral.pdf", Boxplots_Peripheral, width = 18, height = 13)

```

## Boxplots Cytokines None_tumor

```{r}
secretome_data_merged_long_boxplots_None_tumor<- secretome_data_merged_long_boxplots %>%
  filter((Location == "None_tumor"))

Boxplots_None_tumor <- ggplot(secretome_data_merged_long_boxplots_None_tumor, aes(x = Sex, y = Expression, fill = Sex)) +
  geom_boxplot(outlier.shape = NA, width = 0.65, alpha = 0.85) +
  stat_compare_means(
    aes(group = Sex),
    method = "wilcox.test", # or "t.test" depending on your analysis
    label = "p.signif",     # shows stars instead of p-values
    label.x.npc = "center"
  ) +
  geom_jitter(aes(color = Sex),
              width = 0.15, height = 0, size = 1.6, alpha = 0.5, show.legend = FALSE) +
  facet_wrap(~ Cytokine, scales = "free_y") +
  scale_fill_manual(values = c(Male = "#0000FF", Female = "#FF0000")) +
  scale_color_manual(values = c(Male = "#000080", Female = "#D50000")) +
  labs(
    x = "Sex",
    y = "Expression",
    title = "Cytokine Expression by Sex None_tumor",
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.spacing = unit(1, "lines"),
    strip.text = element_text(face = "bold"),
    legend.position = "bottom"
  )

#Print the plot
print(Boxplots_None_tumor)

#Save to file
ggsave("cytokine_boxplots_None_tumor.pdf", Boxplots_None_tumor, width = 18, height = 13)

```

## Boxplots Cytokines Plasma

```{r}
secretome_data_merged_long_boxplots_Plasma<- secretome_data_merged_long_boxplots %>%
  filter((Location == "Plasma"))

Boxplots_Plasma <- ggplot(secretome_data_merged_long_boxplots_Plasma, aes(x = Sex, y = Expression, fill = Sex)) +
  geom_boxplot(outlier.shape = NA, width = 0.65, alpha = 0.85) +
  stat_compare_means(
    aes(group = Sex),
    method = "wilcox.test", # or "t.test" depending on your analysis
    label = "p.signif",     # shows stars instead of p-values
    label.x.npc = "center"
  ) +
  geom_jitter(aes(color = Sex),
              width = 0.15, height = 0, size = 1.6, alpha = 0.5, show.legend = FALSE) +
  facet_wrap(~ Cytokine, scales = "free_y") +
  scale_fill_manual(values = c(Male = "#0000FF", Female = "#FF0000")) +
  scale_color_manual(values = c(Male = "#000080", Female = "#D50000")) +
  labs(
    x = "Sex",
    y = "Expression",
    title = "Cytokine Expression by Sex Plasma",
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.spacing = unit(1, "lines"),
    strip.text = element_text(face = "bold"),
    legend.position = "bottom"
  )

#Print the plot
print(Boxplots_Plasma)

#Save to file
ggsave("cytokine_boxplots_Plasma.pdf", Boxplots_Plasma, width = 18, height = 13)

```

### Plotting selected cytokines

```{r}
secretome_data_merged_long_boxplots_CXCL6<- secretome_data_merged_long_boxplots %>%
  filter((Cytokine == "CXCL6"))

Boxplots_CXCL6 <- ggplot(secretome_data_merged_long_boxplots_CXCL6, aes(x = Sex, y = Expression, fill = Sex)) +
  geom_boxplot(outlier.shape = NA, width = 0.65, alpha = 0.85) +
  stat_compare_means(
    aes(group = Sex),
    method = "wilcox.test", # or "t.test" depending on your analysis
    label = "p.signif",     # shows stars instead of p-values
    label.x.npc = "center"
  ) +
  geom_jitter(aes(color = Sex),
              width = 0.15, height = 0, size = 1.6, alpha = 0.5, show.legend = FALSE) +
  facet_wrap(~ Location, scales = "fixed") +
  scale_fill_manual(values = c(Male = "#0000FF", Female = "#FF0000")) +
  scale_color_manual(values = c(Male = "#000080", Female = "#D50000")) +
  labs(
    x = "Sex",
    y = "Expression",
    title = "CXCL6 Expression by Sex",
  ) +
  theme_minimal(base_size = 30) +
  theme(
    panel.spacing = unit(1, "lines"),
    strip.text = element_text(face = "bold"),
    legend.position = "bottom"
  )

#Print the plot
print(Boxplots_CXCL6)

#Save to file
ggsave("cytokine_boxplots_CXCL6.pdf", Boxplots_CXCL6, width = 11, height = 10)

```

```{r}
secretome_data_merged_long_boxplots_TNFA<- secretome_data_merged_long_boxplots %>%
  filter((Cytokine == "TNFA"))

Boxplots_TNFA <- ggplot(secretome_data_merged_long_boxplots_TNFA, aes(x = Sex, y = Expression, fill = Sex)) +
  geom_boxplot(outlier.shape = NA, width = 0.65, alpha = 0.85) +
  stat_compare_means(
    aes(group = Sex),
    method = "wilcox.test", # or "t.test" depending on your analysis
    label = "p.signif",     # shows stars instead of p-values
    label.x.npc = "center"
  ) +
  geom_jitter(aes(color = Sex),
              width = 0.15, height = 0, size = 1.6, alpha = 0.5, show.legend = FALSE) +
  facet_wrap(~ Location, scales = "fixed") +
  scale_fill_manual(values = c(Male = "#0000FF", Female = "#FF0000")) +
  scale_color_manual(values = c(Male = "#000080", Female = "#D50000")) +
  labs(
    x = "Sex",
    y = "Expression",
    title = "TNFA Expression by Sex",
  ) +
  theme_minimal(base_size = 30) +
  theme(
    panel.spacing = unit(1, "lines"),
    strip.text = element_text(face = "bold"),
    legend.position = "bottom"
  )

#Print the plot
print(Boxplots_TNFA)

#Save to file
ggsave("cytokine_boxplots_TNFA.pdf", Boxplots_TNFA, width = 11, height = 10)

```

```{r}
secretome_data_merged_PDAC_long_summary_heatmap <- rbind(secretome_data_merged_PDAC_females_long_summary,secretome_data_merged_PDAC_males_long_summary)


secretome_data_merged_PDAC_long_summary_heatmap_FC <- secretome_data_merged_PDAC_long_summary_heatmap %>%
  select(Location, Cytokine, Sex, Expression) %>%
  # Go wide to get Male and Female side-by-side
  pivot_wider(names_from = Sex, values_from = Expression) %>%
  # Compute Fold Change and log2FC
  mutate(
    FC_Female_vs_Male = `Female` / `Male`,
    log2FC_Female_vs_Male = log2(`Female`) - log2(`Male`)
  )


```

```{r}
mat <- secretome_data_merged_PDAC_long_summary_heatmap_FC %>%
  select(Location, Cytokine, log2FC_Female_vs_Male) %>%
  pivot_wider(names_from = Location, values_from = log2FC_Female_vs_Male) %>%
  tibble::column_to_rownames("Cytokine") %>%
  as.matrix()

# Choose a symmetric color range around 0
lim <- max(abs(mat), na.rm = TRUE)

Heatmap_Cytokines <- pheatmap(
  mat,
  color = colorRampPalette(c("#0000FF", "white", "#FF0000"))(201),
  breaks = seq(-lim, lim, length.out = 202),
  cluster_rows = TRUE,
  cluster_cols = TRUE,
  main = "log2 Fold Change (Female/Male)",
  na_col = "grey90",         # color for missing tiles
  border_color = NA
)

ggsave("Heatmap_Cytokines.pdf", Heatmap_Cytokines, width = 7, height = 10)
print(Heatmap_Cytokines)
```

## Heatmap females

```{r}
secretome_data_merged_PDAC_females_heatmap <- secretome_data_merged_long_boxplots %>%
  filter((Sex == "Female"))

secretome_data_merged_PDAC_females_heatmap <- secretome_data_merged_PDAC_females_heatmap %>%
  filter((Location != "Plasma"))

secretome_data_merged_PDAC_females_heatmap_summary <- summarySE(secretome_data_merged_PDAC_females_heatmap, measurevar="Log2_FC", groupvars=c("Location", "Cytokine"))

mat_heatmap_female <- secretome_data_merged_PDAC_females_heatmap_summary %>%
  select(Location, Cytokine, Log2_FC) %>%
  pivot_wider(names_from = Location, values_from = Log2_FC) %>%
  tibble::column_to_rownames("Cytokine") %>%
  as.matrix()

# Choose a symmetric color range around 0
lim <- max(abs(mat_heatmap_female), na.rm = TRUE)

Heatmap_Cytokines_females <- pheatmap(
  mat_heatmap_female,
  color = colorRampPalette(c("#0000FF", "white", "#FF0000"))(201),
  breaks = seq(-lim, lim, length.out = 202),
  cluster_rows = FALSE,
  cluster_cols = FALSE,
  main = "log2 Fold Change (Location/Plasma) females",
  na_col = "grey90",         # color for missing tiles
  border_color = NA
)

ggsave("Heatmap_Cytokines_females.pdf", Heatmap_Cytokines_females, width = 4, height = 7)
print(Heatmap_Cytokines_females)
```

```{r}
Heatmap_Cytokines_females_clustered <- pheatmap(
  mat_heatmap_female,
  color = colorRampPalette(c("#0000FF", "white", "#FF0000"))(201),
  breaks = seq(-lim, lim, length.out = 202),
  cluster_rows = TRUE,
  cluster_cols = FALSE,
  main = "log2 Fold Change (Location/Plasma) females",
  na_col = "grey90",         # color for missing tiles
  border_color = NA
)

ggsave("Heatmap_Cytokines_females_clustered.pdf", Heatmap_Cytokines_females_clustered, width = 4, height = 7)
print(Heatmap_Cytokines_females_clustered)
```

## Heatmap male

```{r}
secretome_data_merged_PDAC_males_heatmap <- secretome_data_merged_long_boxplots %>%
  filter((Sex == "Male"))

secretome_data_merged_PDAC_males_heatmap <- secretome_data_merged_PDAC_males_heatmap %>%
  filter((Location != "Plasma"))

secretome_data_merged_PDAC_males_heatmap_summary <- summarySE(secretome_data_merged_PDAC_males_heatmap, measurevar="Log2_FC", groupvars=c("Location", "Cytokine"))

mat_heatmap_female <- secretome_data_merged_PDAC_males_heatmap_summary %>%
  select(Location, Cytokine, Log2_FC) %>%
  pivot_wider(names_from = Location, values_from = Log2_FC) %>%
  tibble::column_to_rownames("Cytokine") %>%
  as.matrix()

# Choose a symmetric color range around 0
lim <- max(abs(mat_heatmap_female), na.rm = TRUE)

Heatmap_Cytokines_males <- pheatmap(
  mat_heatmap_female,
  color = colorRampPalette(c("#0000FF", "white", "#FF0000"))(201),
  breaks = seq(-lim, lim, length.out = 202),
  cluster_rows = FALSE,
  cluster_cols = FALSE,
  main = "log2 Fold Change (Location/Plasma) males",
  na_col = "grey90",         # color for missing tiles
  border_color = NA
)

ggsave("Heatmap_Cytokines_males.pdf", Heatmap_Cytokines_males, width = 4, height = 7)
print(Heatmap_Cytokines_males)
```

```{r}
Heatmap_Cytokines_males_clustered <- pheatmap(
  mat_heatmap_female,
  color = colorRampPalette(c("#0000FF", "white", "#FF0000"))(201),
  breaks = seq(-lim, lim, length.out = 202),
  cluster_rows = TRUE,
  cluster_cols = FALSE,
  main = "log2 Fold Change (Location/Plasma) males",
  na_col = "grey90",         # color for missing tiles
  border_color = NA
)

ggsave("Heatmap_Cytokines_males_clustered.pdf", Heatmap_Cytokines_males_clustered, width = 4, height = 7)
print(Heatmap_Cytokines_males_clustered)
```


## Heatmap females None tumor baseline

```{r}
secretome_data_merged_PDAC_females_heatmap_none_tumor_baseline <- secretome_data_merged_long_boxplots_none_tumor_baseline %>%
  filter((Sex == "Female"))

secretome_data_merged_PDAC_females_heatmap_none_tumor_baseline <- secretome_data_merged_PDAC_females_heatmap_none_tumor_baseline %>%
  filter((Location != "None_tumor"))

secretome_data_merged_PDAC_females_heatmap_none_tumor_baseline_summary <- summarySE(secretome_data_merged_PDAC_females_heatmap_none_tumor_baseline, measurevar="Log2_FC", groupvars=c("Location", "Cytokine"))

mat_heatmap_female <- secretome_data_merged_PDAC_females_heatmap_none_tumor_baseline_summary %>%
  select(Location, Cytokine, Log2_FC) %>%
  pivot_wider(names_from = Location, values_from = Log2_FC) %>%
  tibble::column_to_rownames("Cytokine") %>%
  as.matrix()

# Choose a symmetric color range around 0
lim <- max(abs(mat_heatmap_female), na.rm = TRUE)

Heatmap_Cytokines_females <- pheatmap(
  mat_heatmap_female,
  color = colorRampPalette(c("#0000FF", "white", "#FF0000"))(201),
  breaks = seq(-lim, lim, length.out = 202),
  cluster_rows = FALSE,
  cluster_cols = FALSE,
  main = "log2 Fold Change (Location/None_tumor) females",
  na_col = "grey90",         # color for missing tiles
  border_color = NA
)

ggsave("Heatmap_Cytokines_females_none_tumor_baseline.pdf", Heatmap_Cytokines_females, width = 4, height = 7)
print(Heatmap_Cytokines_females)
```

```{r}
Heatmap_Cytokines_females_clustered <- pheatmap(
  mat_heatmap_female,
  color = colorRampPalette(c("#0000FF", "white", "#FF0000"))(201),
  breaks = seq(-lim, lim, length.out = 202),
  cluster_rows = TRUE,
  cluster_cols = FALSE,
  main = "log2 Fold Change (Location/None_tumor) females",
  na_col = "grey90",         # color for missing tiles
  border_color = NA
)

ggsave("Heatmap_Cytokines_females_clustered_none_tumor_baseline.pdf", Heatmap_Cytokines_females_clustered, width = 4, height = 7)
print(Heatmap_Cytokines_females_clustered)
```

## Heatmap male none tumor baseline

```{r}
secretome_data_merged_PDAC_males_heatmap_none_tumor_baseline <- secretome_data_merged_long_boxplots_none_tumor_baseline %>%
  filter((Sex == "Male"))

secretome_data_merged_PDAC_males_heatmap_none_tumor_baseline <- secretome_data_merged_PDAC_males_heatmap_none_tumor_baseline %>%
  filter((Location != "None_tumor"))

secretome_data_merged_PDAC_males_heatmap_summary_none_tumor_baseline <- summarySE(secretome_data_merged_PDAC_males_heatmap_none_tumor_baseline, measurevar="Log2_FC", groupvars=c("Location", "Cytokine"))

mat_heatmap_female <- secretome_data_merged_PDAC_males_heatmap_summary_none_tumor_baseline %>%
  select(Location, Cytokine, Log2_FC) %>%
  pivot_wider(names_from = Location, values_from = Log2_FC) %>%
  tibble::column_to_rownames("Cytokine") %>%
  as.matrix()

# Choose a symmetric color range around 0
lim <- max(abs(mat_heatmap_female), na.rm = TRUE)

Heatmap_Cytokines_males <- pheatmap(
  mat_heatmap_female,
  color = colorRampPalette(c("#0000FF", "white", "#FF0000"))(201),
  breaks = seq(-lim, lim, length.out = 202),
  cluster_rows = FALSE,
  cluster_cols = FALSE,
  main = "log2 Fold Change (Location/None_tumor) males",
  na_col = "grey90",         # color for missing tiles
  border_color = NA
)

ggsave("Heatmap_Cytokines_males_none_tumor_baseline.pdf", Heatmap_Cytokines_males, width = 4, height = 7)
print(Heatmap_Cytokines_males)
```

```{r}
Heatmap_Cytokines_males_clustered <- pheatmap(
  mat_heatmap_female,
  color = colorRampPalette(c("#0000FF", "white", "#FF0000"))(201),
  breaks = seq(-lim, lim, length.out = 202),
  cluster_rows = TRUE,
  cluster_cols = FALSE,
  main = "log2 Fold Change (Location/None_tumor) males",
  na_col = "grey90",         # color for missing tiles
  border_color = NA
)

ggsave("Heatmap_Cytokines_males_clustered_none_tumor_baseline.pdf", Heatmap_Cytokines_males_clustered, width = 4, height = 7)
print(Heatmap_Cytokines_males_clustered)
```

# Quality of the data set

## Age

```{r}

patient_data_QC <- patient_data
patient_data_QC <- patient_data_QC %>%
  filter((`PC_(no_0_yes__1)` == 1))
patient_data_QC$`Sex_(0_males_1__females)` <- as.character(patient_data_QC$`Sex_(0_males_1__females)`)
patient_data_QC['Sex_(0_males_1__females)'][patient_data_QC['Sex_(0_males_1__females)'] == '0'] <- 'Male'
patient_data_QC['Sex_(0_males_1__females)'][patient_data_QC['Sex_(0_males_1__females)'] == '1'] <- 'Female'

Boxplots_age <- ggplot(patient_data_QC, aes(x = `Sex_(0_males_1__females)`, y = `Age_(at_time_of_OP)`, fill = `Sex_(0_males_1__females)`)) +
  geom_boxplot(outlier.shape = NA, width = 0.65, alpha = 0.85) +
   stat_compare_means(
    aes(group = `Sex_(0_males_1__females)`),
    method = "wilcox.test", # or "t.test" depending on your analysis
    label = "p.signif",     # shows stars instead of p-values
    label.x.npc = "center"
  ) +
  geom_jitter(aes(color = `Sex_(0_males_1__females)`),
              width = 0.15, height = 0, size = 1.6, alpha = 0.5, show.legend = FALSE) +
  #facet_wrap(~ Cytokine, scales = "free_y") +
  scale_fill_manual(values = c(Male = "#0000FF", Female = "#FF0000")) +
  scale_color_manual(values = c(Male = "#000080", Female = "#D50000")) +
  labs(
    x = "Sex",
    y = "Age [Years]",
    title = "Patient age by sex",
  ) +
  theme_minimal(base_size = 12) 
  #theme(
    #panel.spacing = unit(1, "lines"),
    #strip.text = element_text(face = "bold"),
    #legend.position = "bottom"


#Save to file
ggsave("Boxplots_age.pdf", Boxplots_age, width = 8, height = 6)

#Print the plot
print(Boxplots_age)
```

## Differentiation

```{r}
Dotplot_diff <- ggplot(patient_data_QC, aes(x = `Sex_(0_males_1__females)`, y = `diff_(high___1_medium__2_low__3)`, fill = `Sex_(0_males_1__females)`)) +
  geom_dotplot(binaxis = "y", stackdir = "center", binpositions="all")+
  scale_fill_manual(values = c(Male = "#0000FF", Female = "#FF0000"))+
  labs(
    x = "Sex",
    y = "Differentiation grade",
    title = "Differentiation grade PDAC",
  ) +
  theme_minimal(base_size = 12) 

#Save to file
ggsave("Dotplot_diff.pdf", Dotplot_diff, width = 8, height = 6)

#Print the plot
print(Dotplot_diff)
```

## Tumor location

```{r}
Dotplot_location <- ggplot(patient_data_QC, aes(x = `Sex_(0_males_1__females)`, y = `Tumor_location_(0_head_1_body_2_tail)`, fill = `Sex_(0_males_1__females)`)) +
  geom_dotplot(binaxis = "y", stackdir = "center", binpositions="all")+
  scale_fill_manual(values = c(Male = "#0000FF", Female = "#FF0000"))+
  labs(
    x = "Sex",
    y = "Tumor location",
    title = "Tumor location PDAC",
  ) +
  theme_minimal(base_size = 12) 


#Save to file
ggsave("Dotplot_location.pdf", Dotplot_location, width = 8, height = 6)

#Print the plot
print(Dotplot_location)
```

# Table containing the patient cohort

```{r}

patient_data

patient_data_summary <- patient_data %>%
  filter((`PC_(no_0_yes__1)` == 1))
patient_data_summary <- patient_data_summary %>%
  filter(Diagnosis_for_display_purposes == "PDAC")
colnames(patient_data_summary)[2] <- "Age"
colnames(patient_data_summary)[3] <- "Sex"
colnames(patient_data_summary)[5] <- "Tumor_location"
colnames(patient_data_summary)[14] <- "Differentiation"
colnames(patient_data_summary)[24] <- "Pre_op_chemo"

patient_data_summary$Sex <- as.character(patient_data_summary$Sex)
patient_data_summary['Sex'][patient_data_summary['Sex'] == '0'] <- 'Male'
patient_data_summary['Sex'][patient_data_summary['Sex'] == '1'] <- 'Female'

patient_data_summary$Tumor_location <- as.character(patient_data_summary$Tumor_location)
patient_data_summary['Tumor_location'][patient_data_summary['Tumor_location'] == '0'] <- 'Head'
patient_data_summary['Tumor_location'][patient_data_summary['Tumor_location'] == '1'] <- 'Body'
patient_data_summary['Tumor_location'][patient_data_summary['Tumor_location'] == '2'] <- 'Tail'

patient_data_summary$Differentiation <- as.character(patient_data_summary$Differentiation)
patient_data_summary['Differentiation'][patient_data_summary['Differentiation'] == '1'] <- 'High'
patient_data_summary['Differentiation'][patient_data_summary['Differentiation'] == '2'] <- 'Medium'
patient_data_summary['Differentiation'][patient_data_summary['Differentiation'] == '3'] <- 'Low'

patient_data_summary$Pre_op_chemo <- as.character(patient_data_summary$Pre_op_chemo)
patient_data_summary['Pre_op_chemo'][patient_data_summary['Pre_op_chemo'] == '0'] <- 'No'
patient_data_summary['Pre_op_chemo'][patient_data_summary['Pre_op_chemo'] == '1'] <- 'Yes'

selected_cols <- c("Sex", "Age", "Tumor_location", "Differentiation", "Pre_op_chemo")# Filter and store in a new data frame
patient_data_summary_table <- patient_data_summary[selected_cols]

# Create summary table
tbl <- tbl_summary(
  data = patient_data_summary_table,
  by = Sex, # stratify by treatment
  statistic = list(all_continuous() ~ "{mean} ({sd})",
                   all_categorical() ~ "{n} ({p}%)"),
  label = list(Age ~ "Age (years)", Tumor_location ~ "Tumor Location", Differentiation ~ "Differentiation",  Pre_op_chemo ~ "Chemotherapy prior surgery")
) #%>% # Add this for statistical testing 
  #add_p() %>%
  #bold_labels()


gt_tbl <- as_gt(tbl)
gtsave(gt_tbl, "patient_summary.pdf")

```

# Session Info

```{r}
sessionInfo()
```
